### 2018-05-14 Paper/method.tex
..

In this section, we outline the underlying methodology for constructing a 
reduced-order surrogate (ROS) using sensitivity analysis. The term `reduced-order'
in the present context implies that the surrogate is constructed in a 
subspace that sufficiently captures the uncertainty in the model output. 
In the case of PC surrogates used in this work, it implies that the dimensionality
of the polynomial basis functions is effectively reduced.

A possible approach to 
constructing the ROS involves sensitivity analysis of the
uncertain model parameters with respect to a given output, and thereby
disregarding the uncertainty associated with parameters considered as unimportant.
For this purpose, a global sensitivity analysis based on Sobol indices could be pursued.
However, as discussed earlier in sections~\ref{sec:intro} and~\ref{sec:bg}, determining converged 
estimates of Sobol sensitivity indices typically requires tens of thousands
of model evaluations. Consequently, the exercise becomes prohibitive in
situations where the model runs are compute-intensive. Instead, we adopt a novel approach
that focuses on estimating the upper-bound
upper-bound ($\widehat{\mathcal{C}_i\mu_i}$, see Eq.~\ref{eq:bound}) on Sobol total
effect index ($\mathcal{T}(\theta_i)$) for each parameter, $\theta_i$, and using it as a 
metric to screen parameters that are relatively unimportant. Below, we provide a generalized
strategy for parameter screening.

\clearpage
\iffalse
\begin{breakablealgorithm}
  \caption{Parameter screening with DGSMs: A generalized approach.}
  \begin{algorithmic}[1]
    \Procedure{Screening}{}
      \State Draw samples $\{ \bm{\theta}_k \}_{k = 1}^{n_1}$ 
       according to the probability density $\bm{f(\theta)}$.
      \State $N_t = N_t + n_1$
      \State Compute $\bm{g}^k = \nabla_{\bm{\theta}}G(\bm\theta_k)$, 
             $k = 1, \ldots, n_1$.

%%%%%%%%%%%

\begin{breakablealgorithm}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
  \caption{Parameter screening with DGSMs: A generalized approach.}
  \begin{algorithmic}[1]
\Require $\tau > 0$, $\tau_\text{screen} > 0$,
$s_\text{max}$, $\beta > 0$, $\{ \bm{\theta}_k \}_{k = 1}^{n_0}$, $\{ \bm{g}^k \}_{k=1}^{N_\text{total}}$. 
\Ensure $\mathcal{I}_\text{active}$, $\{ \bm{g}^k \}_{k=1}^{N_\text{total}}$, $N_\text{total}$. 
    \Procedure{Screening}{}
    %  \State Draw samples $\{ \bm{\theta}_k \}_{k = 1}^{n_1}$ 
     %  according to the probability density $\bm{f(\theta)}$.
     %   \State $N_\text{total} = N_\text{total} + n_1$.
      \State Compute $\bm{g}^k = \nabla_{\bm{\theta}}G(\bm\theta_k)$, 
             $k = 1, \ldots, n_1$.

@@

    \EndProcedure
  \end{algorithmic}
  \label{alg:screen}
\end{breakablealgorithm}

\bigskip

In case the gradient, $\nabla_{\bm{\theta}}G_k$ cannot be evaluated analytically, we could use finite-difference
to approximate its value as shown in Eq.~\ref{eq:partial} in Algorithm~\ref{alg:screen}. However, as discussed earlier
in section 2.1, estimating $\bm{\mu}$  using $\mathcal{N}$ samples in the full parameter space requires 
model evaluations at model evaluations at $(N_p+1)\mathcal{N}$ samples, compared to only $\mathcal{N}$ samples in 
situations where the derivative is available analytically. Hence, computational effort when using finite differences
is increased by a factor of $(N_p+1)$.

Based on the steps outlined in the above algorithm, it can be said that the set of sample points used for
estimating the screening metric, $\widehat{\mathcal{C}_i\mu_i}$ is enriched until consistency in ranks
between successive iterations as well as a certain degree of convergence in estimates of the screening 
metric for each uncertain parameter is accomplished. However, the amount of computational effort associated
the screening process is limited by the choice of maximum number of iterations Parameter
screening is an integral part of the overall strategy for constructing the ROS as 
depicted below using a flow-diagram.

\bigskip

\begin{figure}[htbp]
\begin{center}

\tikzset{
    arro/.style={{Square[]->}}
}

@@

\end{tikzpicture}
\end{center}

\caption{Flow-diagram outlining the adaptive strategy for constructing reduced-order surrogates.}
\label{fig:flow}
\end{figure}

As illustrated in the above flow-diagram, application of the above methodology to construct an ROS 
relies on the choice of an \textit{appropriate} model output. Since the parameters are screened
using DGSMs, the model output should exhibit a smooth dependence on the
individual uncertain parameters to enable a reasonable approximation of the derivatives in Eq.~\ref{eq:mu}.
Thus, for a given problem involving multiple outputs or model predictions, it is likely that a reduced-order 
surrogate might be possible only for selected outputs. Hence, where possible, it is recommended to a priori assess the
nature of dependence of a given model output on the parameters.

The amount of computational effort associated with the presented methodology can be mainly attributed to
two steps: I. Parameter Screening, and II. Constructing a converged ROS. Computational gains are realized
in situations where constructing the surrogate in the full parameter space is more expensive than the combined cost
associated with these steps. Determining the optimal allocation of computational resources for these steps, however, 
is not possible a priori. Hence, we suggest estimating the DGSMs on a small set of samples (say, $n_1$ = 5) 
initially, and screening with a relatively large value of tolerance, $\tau$ while keeping a track of computational expense.
Tighter tolerances on $\bm{\mu}$ are
gradually imposed until the screening process yields conclusive results pertaining to the possibility of dimension-reduction
or a reasonably converged surrogate is constructed using the available set of model evaluations in the full parameter space. 
Hence, the combined computational cost associated with steps I and II could be less, equal to, or greater than that needed
to construct the surrogate in the full parameter space. To avoid the case of greater computational effort, we suggest 
that the decision to construct the ROS should account for the degree of dimension-reduction predicted at the end of the
screening step ($\alpha$, defined below in Eq.~\ref{eq:alpha}), and the state of convergence of the full-space surrogate. 
Enormous computational savings are in fact possible when dealing with complex, 
compute-intensive models, even if $\alpha$ in the following equation is
not too large.

\be
\alpha = \frac{\mbox{Dimensionality of ROS}}{\mbox{Dimensionality of the full surrogate}}
\label{eq:alpha}
\ee

In order to assess convergence of the ROS, we estimate the leave-one-out cross validation error as
follows:

@@

\noindent where $N_l$ is the number of training points, 
$\widetilde{\mu}~=~\frac{1}{N}\sum\limits_{i=1}^N G(\bm{\theta}_i)$
is the sample mean of the model response, and $ G^{\mbox{\tiny {PC}\textbackslash i}}$
is the PCE surrogate constructed using all but the $i^{\mbox{\tiny{th}}}$ model realization. 
From Eq.~\ref{eq:loo}, it appears that $N$ PCEs are needed to evaluate $\epsilon_{\mbox{\tiny LOO}}$.
However, in practice a modified formulation for $\epsilon_{\mbox{\tiny LOO}}$~\cite{Blatman:2009},
independent of $G^{\mbox{\tiny {PC}\textbackslash i}}$ is used.

Finally, as outlined in Figure~\ref{fig:flow}, the converged ROS is verified for predictive accuracy. 
To this end, we compute a relative L-2 norm of the difference in predictions between the model
and the ROS ($\epsilon_{\mbox{\tiny{L-2}}}$) as follows:

@@

\noindent where $N_v$ is the number of sampling points in the full parameter space at which model
evaluations are available. It must be noted that the computational effort associated with estimating
 $\epsilon_{\mbox{\tiny{L-2}}}$
is negligible since we exploit the available set of model evaluations in the full parameter space, generated
during the screening step. Additionally, probability density function (PDF) based on these full-space 
model evaluations are compared with the PDF generated using a large set of evaluations of the same
output from the ROS.

>>>>

In this section, we outline the underlying framework for ``adaptively" constructing a 
reduced-order surrogate (ROS) using sensitivity analysis.  The proposed methodology
is described as adaptive since the ROS is constructed only in situations where
it is expected to yield computational dividend as discussed further below. 
The term `reduced-order' implies that the surrogate is constructed in a 
subspace that sufficiently captures the uncertainty in the model output. 
In the context of PC surrogates used in this work, it implies that the dimensionality
of the polynomial basis functions is effectively reduced.

Dimension-reduction is guided by global sensitivity analysis (GSA) of the uncertain model parameters. 
A typical approach to GSA involves estimating the Sobol sensitivity indices~\cite{Sobol:2001}
in the case of non-linear mathematical models. However, as discussed earlier in 
sections~\ref{sec:intro} and~\ref{sec:bg}, determining converged 
estimates of Sobol sensitivity indices typically requires tens of thousands
of model evaluations. Consequently, the exercise becomes prohibitive especially in
situations where the model runs are compute-intensive. In the proposed framework,
we adopt a novel approach for constructing an ROS by estimating the upper-bound 
($\widehat{\mathcal{C}_i\mu_i}$, see Eq.~\ref{eq:bound}) on Sobol total
effect index ($\mathcal{T}(\theta_i)$) for each parameter, $\theta_i$, and use it
as a metric to screen parameters that are relatively unimportant. A generalized
methodology for parameter screening is provided below in Algorithm~\ref{alg:screen}.

%%%%%%%%%%% Revised %%%%%%%%%%%
\bigskip
\begin{breakablealgorithm}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
  \caption{Parameter screening with DGSMs: A generalized approach.}
  \begin{algorithmic}[1]
\Require $\tau > 0$, $\tau_\text{screen} > 0$,
$s_\text{max}$, $\beta > 0$, $\{ \bm{\theta}_k \}_{k = 1}^{n_0}$, $\{ \bm{g}^k \}_{k=1}^{N_\text{total}}$. 
\Ensure $\mathcal{I}_\text{active}$, $\{ \bm{g}^k \}_{k=1}^{N_\text{total}}$, $N_\text{total}$. 
    \Procedure{Screening}{}
    %  \State Draw samples $\{ \bm{\theta}_k \}_{k = 1}^{n_1}$ 
     %  according to the probability density $\bm{f(\theta)}$.
     %   \State $N_\text{total} = N_\text{total} + n_1$.
      \State Compute $\bm{g}^k = \nabla_{\bm{\theta}}G(\bm\theta_k)$, 
             $k = 1, \ldots, n_1$.

@@

    \EndProcedure
  \end{algorithmic}
  \label{alg:screen}
\end{breakablealgorithm}
\bigskip

In situations where $\nabla_{\bm{\theta}}G_k$ cannot be evaluated analytically, we could use finite-difference
to approximate its value as shown in Eq.~\ref{eq:partial}. However, as discussed earlier
in section 2.1, estimating $\bm{\mu}$  using $M$ samples in the full parameter space requires 
model evaluations at $(N_p+1)M$ samples, compared to only $M$ samples in 
case the derivative is available analytically. Hence, computational effort when using finite differences
is expected to increase by a factor of $(N_p+1)$.

Based on the steps outlined in the above algorithm, it can be said that the set of sample points used for
estimating the screening metric, $\widehat{\mathcal{C}_i\mu_i}$ is enriched until consistency in ranks
between successive iterations as well as a certain degree of convergence in estimates of the screening 
metric for each uncertain parameter is accomplished. However, the amount of computational effort associated
the screening process is limited by the choice of maximum number of iterations, $s_\text{max}$. Key inputs
include a limiting value of the maximum relative change in the sensitivity measure between succesive iterations
($\tau$), a limiting ratio ($\tau_\text{screen}$) of the sensitivity metric relative to its maximum value,
 a fraction $\beta$ to guide the number of new samples 
($ \lceil \beta n_1 \rceil$ denotes an integer $\leq~\beta n_1$) at each iteration,
a new set of samples ($\{ \bm{\theta}_k \}_{k = 1}^{n_0}$) for the initial screening step in the algorithm,
and the set of all the available gradient evaluations ($\{ \bm{g}^k \}_{k=1}^{N_\text{total}}$). The outputs
include the set of active indices ($\mathcal{I}_\text{active}$) corresponding to the important parameters,
the total number of available model evaluations ($N_\text{total}$), and the enriched set of gradient evaluations
(($\{ \bm{g}^k \}_{k=1}^{N_\text{total}}$)). Parameter screening is an integral part of the overall framework,
illustrated using a flow-diagram in Figure~\ref{fig:flow}.

\begin{figure}[htbp]
\tikzset{
    arro/.style={{Square[]->}}
}

@@

\end{tikzpicture}

\caption{Flow-diagram outlining the adaptive strategy for constructing reduced-order surrogates.}
\label{fig:flow}
\end{figure}
%\clearpage

We begin by setting aside computational resources for constructing a 
validation test suite to be used for assessing the accuracy of the resulting
surrogate. Obviously, the resources allocated for this purpose
depend upon the application as well as total amount of available resource. The applicability
of the proposed framework depends upon the choice of the model output. Since the screening metric
involves computation of partial derivatives in the full parameter space, the output must exhibit
a smooth dependence on each parameter for the approximation to be reasonable. It is therefore likely
that for a given application involving multiple outputs, the ROS can only be constructed for a 
selected few. Hence, where possible, it is prudent to assess the nature of the input-output
relationship for a given model prior to implementing the present framework. The set of required inputs
for parameter screening are initialized, and model evaluations at $n_0$ random samples in the full-space
are computed. These evaluations are used to construct a surrogate in the full-space (FSS) using regression-based
techniques. If the surrogate is found to be sufficiently accurate for the given application, the process
is terminated. However, it is highly unlikely that a full-space surrogate, constructed using a reasonably 
small number of model evaluations would result in a faithful representation of the input-output relationship. 
The available set of model evaluations are utlized and further enhanced during parameter screening as
discussed earlier. At the end of screening, the set of active indices, $\mathcal{I}_\text{active}$ is used
to evaluate $\alpha$ in the following equation:

\be
\alpha = \frac{|\mathcal{I}_\text{active}|}{N_p}
\label{eq:alpha}
\ee

\noindent where $|\mathcal{I}_\text{active}|$ denotes the cardinality of $\mathcal{I}_\text{active}$.
Scope for dimension-reduction increases as $\alpha$ decreases. Hence, if $\alpha$ is
considered to be small and computational gains are expected owing to dimension reduction, the ROS
is constructed and verified for accuracy using a combination of model evaluations used for screening
and those associated with the validation test suite. On the other hand, if $\alpha$ is considerably
large, the set of inputs required for screening are updated as needed, and a new set of $n_0$ samples
and corresponding model evaluations are generated. The FSS is re-constructed using the enriched set
of evaluations and the aforementioned analysis is repeated as illustrated in the flow-diagram.

The amount of computational effort associated with the presented methodology can be mainly attributed to
two steps: I. Parameter Screening, and II. Constructing a converged ROS. Computational gains are realized
in situations where constructing the surrogate in the full parameter space is more expensive than the combined cost
associated with these steps. Determining the optimal allocation of computational resources for these steps, however, 
is not possible a priori. Hence, in the proposed framework, we exploit the set of model evaluations used
in parameter screening to simultaneously construct the FSS while keeping a track of its accuracy using the
validation test suite. This would help address situations where significant dimension
reduction is not possible, and hence, constructing the ROS might result in a computational disadvantage. We
suggest using a small number of samples in the initial screening step (say, $n_0$ = 5) and a relatively large
$\tau$ (say, $\mathcal{O}(10^{-1})$) as a starting point with possible reduction in $\tau$ during
subsequent screenings.

In order to assess convergence of the resulting surrogate, one could estimate the leave-one-out cross 
validation error as follows:

@@

\noindent where $N_l$ is the number of training points, 
$\widetilde{\mu}~=~\frac{1}{N_l}\sum\limits_{i=1}^{N_l} G(\bm{\theta}_i)$
is the sample mean of the model response, and $ G^{\mbox{\tiny {PC}\textbackslash i}}$
is the PC surrogate constructed using all but the $i^{\mbox{\tiny{th}}}$ model realization. 
From Eq.~\ref{eq:loo}, it appears that $N_l$ PCEs are needed to evaluate $\epsilon_{\mbox{\tiny LOO}}$.
However, in practice a modified formulation for $\epsilon_{\mbox{\tiny LOO}}$~\cite{Blatman:2009},
independent of $G^{\mbox{\tiny {PC}\textbackslash i}}$ is used. Accuracy of the
surrogate could be assessed by evaluating the relative L-2 norm of the difference in predictions
between the model and the surrogate ($\epsilon_{\mbox{\tiny{L-2}}}$), as follows:

@@

\noindent where $N_v$ is the number of sampling points in the full parameter space at which model
evaluations are available. Accuracy of the surrogate could be further investigated by comparing
probability density functions (PDFs) of the model output based on model evaluations in the full
parameter space and surrogate predictions corresponding to a large number of samples.

%\newpage
\iffalse
\begin{breakablealgorithm}
  \caption{Parameter screening with DGSMs: A generalized approach.}
  \begin{algorithmic}[1]
    \Procedure{Screening}{}
      \State Draw samples $\{ \bm{\theta}_k \}_{k = 1}^{n_1}$ 
       according to the probability density $\bm{f(\theta)}$.
      \State $N_t = N_t + n_1$
      \State Compute $\bm{g}^k = \nabla_{\bm{\theta}}G(\bm\theta_k)$, 
             $k = 1, \ldots, n_1$.

%\newpage
